## Data exploration

### Including packages

```{r}
library(data.table)
library(magrittr)
library(ggplot2)
library(tidyverse)
library(patchwork)
```
### Fetch data
#### Read CSV File
```{r}
survey_df <- read.csv("from_yao/SimulatedUsers-Final_August4.csv")
# survey_df
```

#### Convert to `data.table`
```{r}
dt <- data.table(survey_df)
survey_dt <- dt[-(1:2)]
# summary(survey_dt)
``` 

### Data Wrangling
#### StartDate, EndDate and RecordedDate are dates
```{r}
survey_dt$StartDate <- as.POSIXct(survey_dt$StartDate, format="%Y-%m-%d %H:%M:%S")
survey_dt$StartDate <- as.POSIXct(survey_dt$EndDate, format="%Y-%m-%d %H:%M:%S")
survey_dt$StartDate <- as.POSIXct(survey_dt$RecordedDate, format="%Y-%m-%d %H:%M:%S")
```

#### Study statistics
```{r}
survey_dt$Duration..in.seconds. <- as.integer(survey_dt$Duration..in.seconds.)
```

#### Remove columns
```{r}
survey_dt[, c("X","Unnamed..0"):=NULL]
```

#### Demographic Questions
```{r}
survey_dt$D1 <- as.integer(survey_dt$D1)
setnames(survey_dt, "D1", "Age")

sex_options <- c("F", "M", "X")
survey_dt$D2 <- factor(survey_dt$D2, levels = c("1", "2", "3"), labels = sex_options)
setnames(survey_dt, "D2", "Sex")

# Mapping:
# 2 Student
# 3 Freelancer
# 4 Unemployed/Other
# 5 Part-time
# 6 Full-time
# 8 Unemployed/Other
employment_options <- c()
#survey_dt$D2 <- as.factor(survey_dt$D2)
survey_dt$D3 <- as.factor(survey_dt$D3)
setnames(survey_dt, "D3", "Employment")

```

##### Previous Experience
```{r}
tmp <- survey_dt$D4

mc_binary_encode <- function(s) { 
      no_exp <- grepl('1', s, fixed=TRUE)
      user <- grepl('2', s, fixed=TRUE)
      dev <- grepl('3', s, fixed=TRUE)
      
      res <- paste(no_exp, user, dev)
      
      return(res)
}

survey_dt$D4 <- mc_binary_encode(survey_dt$D4)
survey_dt <- survey_dt %>% separate(D4, c('No_Experience', 'AI_User', 'AI_Dev'))

```

#### Single choice predictions - Birds
```{r}
bird_options <- c("Crested Auklet", "Least Auklet", "Parakeet Auklet", "Rhinoceros Auklet")
numbercol_to_birdnames <- function(s) { factor(s, levels = c("1", "2", "3", "4"), labels = bird_options) }
survey_dt <- survey_dt %>% mutate(across(c(Q1, Q2, Q3, Q4, Q5, Q6, Q7, Q8, Q9, Q10, Q11, Q12, Q13, Q14, Q15, Q16, Q17, Q18, Q19, Q20), numbercol_to_birdnames))
```

#### Single Choice predictions - Street signs
```{r}
sign_options <- c("Left-bend", "Right-bend", "Ahead-or-left", "Ahead-or-right")
numbercol_to_signnames <- function(s) { factor(s, levels = c("1", "2", "3", "4"), labels = sign_options) }
survey_dt <- survey_dt %>% mutate(across(c(Q1.1, Q2.1, Q3.1, Q4.1, Q5.1, Q6.1, Q7.1, Q8.1, Q9.1, Q10.1, Q11.1, Q12.1, Q13.1, Q14.1, Q15.1, Q16.1, Q17.1, Q18.1, Q19.1, Q20.1), numbercol_to_signnames))
```

### Visualizations
#### Demographics
##### Ages
```{r}
ggplot(survey_dt, aes(x = Age)) + geom_bar()
```
##### Sex
```{r}
ggplot(survey_dt, aes(x = Sex)) + geom_bar()
```

##### Employment
```{r}
ggplot(survey_dt, aes(x = Employment)) + geom_bar()

# Mapping:
# 2 Student
# 3 Freelancer
# 4 Unemployed/Other
# 5 Part-time
# 6 Full-time
# 8 Unemployed/Other

```

##### Prior Experience
```{r}
ggplot(survey_dt, aes(x = AI_User, y = AI_Dev)) + geom_point() + geom_jitter()
```

#### Bird Species
##### Guessing Predictions
```{r}
# TODO
```

##### Subjective Understanding
```{r}
sub1_p <- ggplot(survey_dt, aes(x = Q_1)) + geom_bar() + ggtitle("I understood the explanations within\nthe context of this study") + theme(plot.title = element_text(size=8))
sub2_p <-ggplot(survey_dt, aes(x = Q_2)) + geom_bar() + ggtitle("The explanations provided enough\ninformation for me to understand\nhow the model arrived at its decision") + theme(plot.title = element_text(size=8))
sub3_p <-ggplot(survey_dt, aes(x = Q_3)) + geom_bar() + ggtitle("I think that most people would learn\nto understand the explanations very\nquickly") + theme(plot.title = element_text(size=8))
sub4_p <-ggplot(survey_dt, aes(x = Q_8)) + geom_bar() + ggtitle("I would like to have more examples\nto understand the machine’s\nreasoning and how the model\narrived at its decision.") + theme(plot.title = element_text(size=8))
sub5_p <-ggplot(survey_dt, aes(x = Q_5)) + geom_bar() + ggtitle("The explanations were useful and\nhelped me understand the model's\nreasoning.") + theme(plot.title = element_text(size=8))
sub6_p <-ggplot(survey_dt, aes(x = Q_6)) + geom_bar() + ggtitle("I believe that I could provide an\nexplanation similar to the model's\nexplanation for a new image.") + theme(plot.title = element_text(size=8))
sub1_p + sub2_p + sub3_p + sub4_p + sub5_p + sub6_p + plot_layout(ncol = 3)
```

#### Street Signs
##### Guessing Predictions
```{r}
# TODO
```

##### Subjective Understanding
```{r}
sub1.1_p <- ggplot(survey_dt, aes(x = Q_1.1)) + geom_bar() + ggtitle("I understood the explanations within\nthe context of this study") + theme(plot.title = element_text(size=8))
sub2.1_p <-ggplot(survey_dt, aes(x = Q_2.1)) + geom_bar() + ggtitle("The explanations provided enough\ninformation for me to understand\nhow the model arrived at its decision") + theme(plot.title = element_text(size=8))
sub3.1_p <-ggplot(survey_dt, aes(x = Q_3.1)) + geom_bar() + ggtitle("I think that most people would learn\nto understand the explanations very\nquickly") + theme(plot.title = element_text(size=8))
sub4.1_p <-ggplot(survey_dt, aes(x = Q_8.1)) + geom_bar() + ggtitle("I would like to have more examples\nto understand the machine’s\nreasoning and how the model\narrived at its decision.") + theme(plot.title = element_text(size=8))
sub5.1_p <-ggplot(survey_dt, aes(x = Q_5.1)) + geom_bar() + ggtitle("The explanations were useful and\nhelped me understand the model's\nreasoning.") + theme(plot.title = element_text(size=8))
sub6.1_p <-ggplot(survey_dt, aes(x = Q_6.1)) + geom_bar() + ggtitle("I believe that I could provide an\nexplanation similar to the model's\nexplanation for a new image.") + theme(plot.title = element_text(size=8))
sub1.1_p + sub2.1_p + sub3.1_p + sub4.1_p + sub5.1_p + sub6.1_p + plot_layout(ncol = 3)
```

#### Export
```{r}
write.csv(survey_dt, "cleaned_simulatedusers.csv")
```